---
title: "Chapter 3 Notes"
author: "Joshua Susanto"
date: "2023-02-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


Chapter 3 Notes

# Main Tidyverse Packages
tibble: A package to introduce a more efficient data frame.
readr: Functions to read rectangular data.
tidyr: Functions to help make data “tidy.”
dplyr: Functions for data manipulation.
purrr: Functions to enhance R’s functional programming (i.e., vectorization).
stringr: Functions for string (character) manipulation.
forcats: Functions for factors (categorical variables).
ggplot2: A graphics system based on the “Grammar of Graphics.”


# %>% Operator
- f(x) can be written as x %>% f()
- f(x,y) can be written as x %>% f(y)
- leads input into the first argument of the function

x %>% f(y) %>% g(z) -> g(f(x,y),z)

# Argument Placeholder
The pipe operator can also be used as an argument placeholder by using a .
on the righthand side to represent the object on the lefthand side

f(x,y) can be written as y %>% f(x,.)

ex: trees %>% lm(Volume ~ Height,data=.)


# Tibbles
- Tibbles are one of the central data structures used in the tidyverse
- Tibbles are data frames with the added class tbl_df that makes the way R prints the object more readable 
- The tibble() function can create tibbles, the same way that data.frame() does.
- The as_tibble() function coerces lists and matrices into tibbles.

```{r}
# different ways to create tibble

# read_csv -> separate by newline, first row are the column names, created geometrically. don't need commas to separate newline
read_csv("X, Y, Z
    1, 1, 1
    2, 3, 4")

# tibble() -> give column names with their values, separate each column with a comma as a separate argument
tibble(
   X = c(1, 2),
   Y = c(1, 3),
   Z = c(1, 4)
  )

# same as first, but putting newline manually
read_csv("X, Y, Z \n 1, 1, 1\n 2, 3, 4")

# a row wise version of tibble(), similar geometrically to read_csv, put tildas before column names, need commas to separate new line
tribble(
    ~X, ~Y, ~Z,
   #---/---/---,
    1,  1,  1,
    2,  3,  4,
)

```


# Tibbles Versus Data Frames
- Column data is not coerced, a character vector is not coerced into a factor.
- Subsetting a column from a tibble using the single bracket [,j] always returns a tibble vs a vector
- The $ operator does not allow partial name matching the way it does
for data frames

- tibble always prints first 10 rows and all the columns that fit on the screen
- width argument to specify how many columns to put
- width = Inf prints all columns but puts the ones that dont fit on the next line

```{r}
data(diamonds)
diamonds[['carat']] # can subset columns with double bracket and name in quotes, still returns a vector rather than a tibble
diamonds[, 'carat'] # can also use dimensional ss
diamonds[,1] # positional ss
diamonds$carat # $ operator also returns a vector still (similar to double bracket)
```

# readr

read_csv() reads comma delimited files.
read_csv2() reads semicolon delimited files.
read_tsv() reads tab delimited files.
read_delim() reads files with any delimiter.
read_fwf() reads fixed width files.
read_table() reads files with white space separators.


Optional Arguments in read_csv()
- skip = n to skip the first n lines (e.g., if there is metadata at the top of the file).
- comment = "#" to drop all lines starting with the # character.
- col_names = FALSE to read files without column names. The columns will be labeled sequentially from X1 to Xn (for n columns). Alternatively, input a character vector in col_names to specify column names.
- na argument to specify the character(s) that represent missing values in the file.

# Tidy Data
A dataset is called tidy if the following three rules are satisfied:
Each variable must have its own column.
Each observation must have its own row.
Each value must have its own cell.

The first step in tidying data is to figure out what the variables and
observations are.

The main functions in the tidyr package help make data tidy.
Once the variables and observations are specified, there are several functions that address common issues with messy data:

- pivot_longer is used when one variable is spread across multiple
columns.
- pivot_wider is used when one observation is scattered across
multiple rows.
- separate() is used when cells contain multiple values (from different
variables).
- unite() is used when a single variable is spread across multiple
columns.


#dplyr
