---
title: "Chapter 5 Notes"
author: "Joshua Susanto"
date: "2023-02-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(sloop)
library(tidyverse)
```


## CHAPTER 5: OOP IN R

Object-oriented programming (OOP) is a style of programming that focuses on defining different types of objects and functions that can be applied to those object types. Central to any object-oriented system are the concepts of class and method.
- A class is a definition of an object. Typically a class contains several fields that are used to hold class-specific information.
- A method is a function for a specific class.
Intuitively, a class defines what an object is, and methods describe what
that object can do

# Inheritance
Another common feature in object-oriented programming languages is the concept of inheritance. Classes are usually organized in a hierarchy, so if a method does not exist for a child, then its parent’s method is used. We say the child inherits behavior from the parent.

For example, an ordered factor inherits from a regular factor, and a tibble
inherits from a data frame.


# Polymorphism
The primary reason to use OOP is polymorphism (which means “many shapes”), which refers to the ability to have a common interface for different types of inputs
- Polymorphic functions are ones which behave differently for different types of input objects.
- examples: print(), summary(), and plot().

```{r}
data("diamonds")
summary(diamonds$depth)
summary(diamonds$cut)
```

# Three OOP Systems
Base R has three object-oriented systems (plus the base types).
S3 is R’s first, simplest, and most flexible object-oriented system. S3 is
the only object-oriented system used in the base and stats packages,
and it is the most commonly used system in CRAN packages.
S4 is a formalization of S3 that has much stricter implementation for
defining classes, methods, and generic functions. S4 is implemented in
the base methods package.
RC, short for reference classes, is a very different system from S3
and S4. RC implements “message-passing” OOP, so methods belong
to classes, not functions.
We will focus on the S3 system.


# Base Types
There is one other system that is not quite object-oriented, but is important to mention:

Base types are the internal C-level types that are the basis of the other object-oriented systems.
Base types do not form an OOP system, but they provide the building
blocks for the other OOP systems.

