---
title: "102adiscweek5"
author: "Joshua Susanto"
date: "2023-02-08"
output: html_document
---


3. Q
- hw [let grader know]
- regex
- any [have to use with filter and groupby]
- trim 
- join [goes to next column]
- unique key (unique)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```


```{r}
ingredients <- read_csv('ingredient.csv')
ingredients
```


a. Write R code to return the food items required to cook the recipe “Beef Burger”.
```{r}
ingredients %>% filter(recipe == 'Beef Burger') %>% select(food_item)
```

b. Write R code to return all recipes containing both bread and tomato products with their
weights on these two ingredients.

use any because of the group by (IDGI)
```{r}
ingredients %>% group_by(recipe) %>% filter(any(food_item == 'Bread') & any(food_item == 'Tomato' | food_item == 'Tomato Sauce'))
?substr
```

```{r}
ingredients %>% group_by(recipe) %>% filter(any(food_item == 'Bread') & any(substr(food_item, 1, 6) == 'Tomato')) %>% filter(food_item == 'Bread' | substr(food_item, 1, 6) == 'Tomato') %>% group_by(recipe) %>% summarize(total_weight = sum(`weight (oz)`))
```

c. Write R code to return the food items and their prices for all items of meat products sold at
“MyMart” or “Coco Mart”.

```{r}
food_item <- read_csv('food_item.csv')
food_item
```

```{r}
stock <- read_csv('stock.csv')
stock
```

```{r}
?trimws
unique(stock$food_item)
unique(food_item$item)
```

#use vector to select multiple columns
```{r}
stock %>% left_join(food_item, by = c('food_item' = 'item')) %>% filter(type == 'Meat', (shop == 'MyMart' | shop == 'Coco Mart')) %>% select(c(food_item, shop, `price (US dollars per lb)`))
```


d. Write R code to return the recipes with the total calories that contain “Wheat product”.

```{r}
recipe <- read_csv('recipe.csv')
recipe
```

```{r}
ingredients 
```

