---
title: "102a Discussion 9"
author: "Joshua Susanto"
date: "2023-03-08"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

1: Dealing with Large Numbers
To simulate how the computer handles the computation with large floating point numbers we
define pqnumber objects to keep large numbers. This object should have four components. The
first one is either the integer +1 or the integer −1. It gives the sign of the number. The second
and third are p and q. And the fourth component is a vector of p + q + 1 integers between zero
and nine. For example, we can use the following object x to keep the number 87654.321.

```{r}
y <- structure(list(sign = 1, p = 3, q = 4, nums = 1:8), class = "pqnumber")
# nums give the numbers from right to left
# p gives the place of the decimal
# q gives how many to the left of decimal (starts at 0)
y
```


(a) Write a constructor function, an appropriate predicate function, appropriate coercion
functions, and a useful print() method. This question accounts for 25% of this assignment.

• The constructor takes the four arguments: sign, p, q, and nums. Then check if the
arguments satisfy all requirements for a pqnumber object. If not, stop with an
appropriate error message. If yes, return the object.
• A predicate is a function that returns a single TRUE or FALSE, like is.data.frame(), or
is.factor(). Your predicate function should be is pqnumber() and should behave as
expected.
• A useful print() method should allow users to print a pqnumber object with the
decimal representation defined as follows:

CONSTRUCTOR
```{r}
# needs flow chart and test cases
pqnumber <- function(sign, p, q, nums) {
  if(abs(sign) != 1) {
    stop('Sign should be +/- 1')
  }
  # need to check that p and q are nonnegative integers (whole numbers including 0)
  # nums should be exactly length p + q + 1
  # nums is a vector of digits
  structure(list(sign = as.integer(sign), p = as.integer(p), q = as.integer(q), nums = as.integer(nums)), class = 'pqnumber')
  
}

# TEST CASES ALL TRUE
x <- pqnumber(1, 2, 0, c(4, 1, 3))
x$sign == 1
x$p == 2
x$ q== 0
all(x$nums == c(4, 1, 3)) 
```


```{r}
c(1,2) == c(2,2) # notice the behavior of vector comparison, why we use all and any
```

```{r}
y <- pqnumber(0.5, 2, 0, c(4, 1, 3)) #should be an error
```


PREDICATE

```{r}
# flow charts and test cases
# cant just check if the class is pqnumber
# an object can have multiple classes

is_pqnumber <- function(x) {
  any(class(x) == 'pqnumber')
}
```


```{r}
is_pqnumber(x) # true
is_prnumber(3) # false
is_pqnumber(matrix(0,3,3)) #false
```


PRINT

```{r}
print.pqnumber <- function(x, DEC = FALSE) {
  if (DEC) {
    if (x$sign == -1) {
      sbit <- '-'
    } else {
      sbit <- ''
    }
    decimal <- paste(rev(x$nums[0:x$p]), collapse = '')
    number <- paste(rev(x$nums[(x$p + 1):(x$p+x$q+1)]), collapse = '')
    print_str <- paste(sbit, number, '.', decimal, sep = '')
  } else {
    print_str <- paste0("sign = ", x$sign, '\np = ', x$p, '\nq = ', x$q, '\nnums = ', paste(x$nums, collapse = ' '))
  }
  cat(paste0(print_str, '\n'))
}
```

```{r}
print.pqnumber(x)
print.pqnumber(y, DEC = TRUE)
```


AS PQNUMBER

```{r}
as_pqnumber <- function(x, p, q) {
  UseMethod('as_pqnumber')
}

as_pqnumber.numeric <- function(x, p, q) {
  nums <- (abs(x) * 10^seq(p, -q)) %% 10 %/% 1
  sgn <- if(x == 0) 1 else base::sign(x)
  pqnumber(sgn, p, q, nums)
}

as_pqnumber(3.14, 2, 0)
```

AS NUMERIC GENERUC AS NUMERIC.PQNUMBER : METHOD

```{r}
as_numeric <- function(x) {
  UseMethod('as_numeric')
}
as_numeric.pqnumber <- function(x) {
  #converts pqnumber to numeric
  #Args:
  #x: pqnumber size 1
  #Return:
  #numeric size 1
  dec_num <- 0
  for(i in seq_along(x$nums)) {
    dec_num <- dec_num + (x$nums[i] * 10^(i-1))
  }
  exp <- -1 * x$p
  dec_num <- dec_num * x$sign * (10^exp)
  dec_num
}

as_numeric.pqnumber(x)
as_numeric(y)
```


 B)
 
Addition and subtraction

```{r}
# note that borrowing and carry over are going to be one function
abs_gtr <- function(a, b) {
  if (a$q > b$q) {
    b <- pqnumber(b$sign, b$p, a$q, c(b$nums, rep(0L, (a$q - b$q))))
  } else if (b$q > a$q) {
    a <- pqnumber(a$sign, a$p, b$q, c(a$nums, rep(0L, (b$q - a$q))))
  }
  for (i in 1:(a$q + 1)) {
    if (rev(a$nums)[i] > rev(b$nums)[i]) {
      return(TRUE)
    } else if (rev(a$nums)[i] < rev(b$nums)[i]) {
      return(FALSE)
    }
  }
}

carry_over <- function(z) {
  n <- length(z)
  carry <- 0
  for (i in 1:n) {
    zi <- z[i] + carry
    z[i] <- zi %% 10
    carry <- zi %/% 10
  }
  if (carry != 0) {
    z[n+1] <- carry
  }
  z
}

add <- function(x, y) {
  # error handle
  max_p <- max(x$p, y$p)
  max_q <- max(x$q, y$q)
  n <- max_p + max_q + 1
  z <- rep(0L, n)
  if(x$sign == y$sign) {
    x_vals <- x$nums
    y_vals <- y$nums
    sgn <- x$sign
  } else {
    if(abs_gtr(x,y)) {
      x_vals <- x$nums
      y_vals <- -y$nums
      sgn <- x$sign
    } else {
      x_vals <- -x$nums
      y_vals <- y$nums
      sgn <- y$sign
    }
  }
  z[(1 + max_p - x$p):(1 + max_p + x$q)] <- x_vals
  z[(1 + max_p - y$p):(1 + max_p + y$q)] <- z[(1 + max_p - y$p):(1 + max_p + y$q)] + y_vals
  z <- carry_over(z)
  digit_offset <- length(z) - n
  pqnumber(sgn, max_p, max_q + digit_offset, z)
}


subtract <- function(x, y) {
  y$sign <- y$sign * -1
  add(x, y)
}
``` 
 
```{r}
add(x,x)
add(y, y)
add(x, y)
subtract(y, x)
subtract(x, y)
neg <- pqnumber(-1, 3, 4, c(1,2,3,4, 0,0,0,0))
pos <-pqnumber(1, 3, 3, c(1,2,3,0, 1,0,0))
add(pos, neg)
add(neg, pos)
subtract(pos, neg)
subtract(neg, pos)
```
 
```{r}
tes1 <-  pqnumber(1, 3, 4, c(1,2,3,4, 0,0,0,0))
print.pqnumber(tes1, DEC = TRUE)
tes2 <-  pqnumber(1, 3, 3, c(1,2,3,0, 1,0,0))
print.pqnumber(tes2, DEC = TRUE)
tes3 <- pqnumber(tes2$sign, tes2$p, tes1$q, c(tes2$nums, rep(0, (tes1$q - tes2$q))))
print.pqnumber(tes3, DEC = TRUE)
abs_gtr(x, y)
abs_gtr(y, x)
abs_gtr(tes1, tes2)
```
 
 MULTIPLY
 
```{r}
multiply <- function(x, y) {
  n <- x$p + y$p + x$q + y$q + 1
  z <- rep(0L, n)
  for (r in 1:(1 + y$p + y$q)) {
    x_leftover <- x$p + x$q 
    z[r:(r + x_leftover)] <- z[r:(r + x_leftover)] + (x$nums * y$nums[r])
  }
  z <- carry_over(z)
  digit_offset <- length(z) - n
  sgn <- x$sign * y$sign
  p_new <- x$p + y$p
  q_new <- length(z) - p_new - 1
  pqnumber(sgn, p_new, q_new, z)
}
multiply(pqnumber(1, 2, 0, c(4, 1, 3)), pqnumber(1, 2, 0, c(4, 1, 3)))
```
 
 