---
title: "102adiscweek5"
author: "Joshua Susanto"
date: "2023-02-08"
output: html_document
---

Q
- 1.1d
- 2.2a
- 2.3a

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
```


```{r}
ingredients <- read_csv('ingredient.csv')
ingredients
```


a. Write R code to return the food items required to cook the recipe “Beef Burger”.
```{r}
ingredients %>% filter(recipe == 'Beef Burger') %>% select(food_item)
```

b. Write R code to return all recipes containing both bread and tomato products with their
weights on these two ingredients.

use any because of the group by
```{r}
ingredients %>% group_by(recipe) %>% filter(any(food_item == 'Bread') & any(food_item == 'Tomato' | food_item == 'Tomato Sauce'))
?substr
```

```{r}
ingredients %>% group_by(recipe) %>% filter(any(food_item == 'Bread') & any(substr(food_item, 1, 6) == 'Tomato')) %>% filter(food_item == 'Bread' | substr(food_item, 1, 6) == 'Tomato') %>% summarize(total_weight = sum(`weight (oz)`))
```

c. Write R code to return the food items and their prices for all items of meat products sold at
“MyMart” or “Coco Mart”.

```{r}
food_item <- read_csv('food_item.csv')
food_item
stock <- read_csv('stock.csv')
stock
unique(stock$food_item)
unique(food_item$item)
```

#use vector to select multiple columns
```{r}
stock %>% left_join(food_item, by = c('food_item' = 'item')) # note the behavior of left join when there are multiple keys of the same name

stock %>% left_join(food_item, by = c('food_item' = 'item')) %>% filter(type == 'Meat', (shop == 'MyMart' | shop == 'Coco Mart')) %>% select(c(food_item, shop, `price (US dollars per lb)`))
```


d. Write R code to return the recipes with the total calories that contain “Wheat product”.

total calories of recipes that contain wheat??

```{r}
recipe <- read_csv('recipe.csv')
recipe
```

```{r}
ingredients 
food_item
```
```{r}
ingredients %>% left_join(food_item, by = c('food_item' = 'item')) %>% group_by(recipe) %>% summarize(total_cal = sum(calories))


has_wheat <- ingredients %>% left_join(food_item, by = c('food_item' = 'item')) %>% group_by(recipe) %>% filter(type == 'Wheat product') %>% select %>% pull()

ingredients %>% left_join(food_item, by = c('food_item' = 'item'))
```

```{r}
ingredients %>% left_join(food_item, by = c('food_item' = 'item')) %>% filter(recipe %in% has_wheat) %>% group_by(recipe) %>% summarize(total_cal = sum(calories))  
```


2: Regular Expression Exercises
• The exercises focus on using regular expressions and the stringer functions.
• Don’t use any control statements.
• Please find and download the corresponding files on the course website to complete the
tasks below.
• For each pattern, store the pattern as a variable named in the way pat_x_y where the x is the number of the exercise and the y is the letter of the sub-exercise.
• You should save these patterns in your .R file.
• You may demonstrate your patterns with any appropriate stringer functions.
• The accuracy of your patterns account for 60% of this assignment.


1. Find names.txt and write regex patterns to answer following questions.
a. Find all usernames that contain at least one numeric character.

```{r}
names <- read.table(file = 'names.txt', sep = '\n')
names <- names$V1
names
pat_1_a <- '.*[0-9].*'
str_extract_all(names, pat_1_a)
```

b. Find all usernames that are exactly four characters long and consist only of alphabetic

```{r}
pat_1_b <- '\\A[[:alpha:]][[:alpha:]][[:alpha:]][[:alpha:]]\\Z'
str_match_all(names, pat_1_b)
```


c. Find all usernames following the conventional way of name format, i.e., the “given name” goes first, and the “family” name last, with any other names in-between. The names are separated by a single white space and each name should be uppercase letter followed by one or more lowercase letters.

```{r}
pat_1_c <- '\\A[A-Z][[:alpha:]]+ ([A-Z][[:alpha:]]+ )*[A-Z][[:alpha:]]+\\Z'
str_match_all(names, pat_1_c)
str_match_all('Joshua  Susanto', pat_1_c)
str_match_all('Joshua Susanto', pat_1_c)
str_match_all('Joshua Joshj Susanto', pat_1_c)
str_match_all('Joshua Joshj meow Susanto', pat_1_c)
str_match_all('Joshua Joshj Joosh Susanto', pat_1_c)
str_match_all('Joshua Joshj  Joosh Susanto', pat_1_c)
```


2. Find cards.txt and write the regex patterns to identify all the valid card numbers with the formatting rules below

• A Master card number begins with a 5 and it is exactly 16 digits long.
• A Visa card number begins with a 4 and it is between 13 and 16 digits long.

a. Write a regex pattern to match valid Master card number and print all the valid
numbers, grouped into sets of 4 separated by a single space.

** Does it need to be separated by a space ?
```{r}
cards <- read.table(file = 'cards.txt', sep = '\n')
cards <- cards$V1
cards
pat_2_a <- '\\A(5[0-9]{3})[ ]*([0-9]{4})[ ]*([0-9]{4})[ ]*([0-9]{4})\\Z'
str_match_all(cards, pat_2_a)
```



b. Write a regex pattern to match valid Visa card number and print all the valid numbers,
grouped into sets of 4 separated by a single space.
Note: You may use more than one stringer function for both questions above.


```{r}
pat_2_b <- '\\A(4[0-9]+)'
```

3. Find the passwords.txt and write regex patterns to answer the following questions

a. Write a regex pattern to identify the passwords that satisfies the requirements below.
• Minimum 8 characters
• Must contain at least one letter
• Must contain at least one digit
- can the other characters be other things?
```{r}
passwords <- read.table(file = 'passwords.txt', sep = '\n')
passwords <- passwords$V1
passwords
pat_3_a <- '\\A(?=.*[[:alpha:]])(?=.*[0-9])[[:alnum:]]{8,}\\Z' #how does this work for ordering (1brownie)
str_match_all(passwords, pat_3_a)
```


b. Write a regex pattern to identify the passwords that satisfies the requirements below.
• Minimum 8 characters
• Must contain at least one uppercase character
• Must contain at least one lowercase character
• Must contain at least one digit
```{r}
pat_3_b <- '\\A(?=.*[A-Z])(?=.*[a-z])(?=.*[0-9])[[:alnum:]]{8,}\\Z' #how does this work for ordering (1brownie), does this act as an if statement
str_match_all(passwords, pat_3_b)
```



4. Please load wordlists.RData to R and write regular expression patterns which will match all
of the values in x and none of the values in y.
a. Ranges

- y only has first words start with g and after other than 6
- of these 6 one contains m, u, j, y, r which are not in x
- 5 of these 6 have letters on the fourth slot not in x
- the last 1 has the second letter as o which is not in x

```{r}
load('wordlists.RData')
range_x <- unlist(wordlists$Ranges[1])
range_y <- unlist(wordlists$Ranges[2])
pat_4_a <- '\\A[a-f][^o].[^mujyh].*'
str_match_all(range_x, pat_4_a)
str_match_all(range_y, pat_4_a)
```

b. Backrefs
```{r}

```

