---
title: "102a Homework 3"
output: html_document
date: "2023-02-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("405568250_stats102a_hw3.R")
library(tidyverse)
```


## Question 1

a. Write R code to return the food items required to cook the recipe “Beef Burger”.

```{r}
ingredients <- read_csv('ingredient.csv')
ingredients %>% filter(recipe == 'Beef Burger') %>% select(food_item)
```

b. Write R code to return all recipes containing both bread and tomato products with their
weights on these two ingredients.

```{r}
ingredients %>% 
  group_by(recipe) %>% 
  filter(any(food_item == 'Bread') & any(substr(food_item, 1, 6) == 'Tomato')) %>%
  filter(food_item == 'Bread' | substr(food_item, 1, 6) == 'Tomato') %>%
  summarize(total_weight_oz = sum(`weight (oz)`))
```

c. Write R code to return the food items and their prices for all items of meat products sold at “MyMart” or “Coco Mart”.

```{r}
food_item <- read_csv('food_item.csv')
stock <- read_csv('stock.csv')
```


```{r}
stock %>% 
  left_join(food_item, by = c('food_item' = 'item')) %>% 
  filter(type == 'Meat', (shop == 'MyMart' | shop == 'Coco Mart')) %>% 
  select(c(food_item, shop, `price (US dollars per lb)`))
```

d. Write R code to return the recipes with the total calories that contain “Wheat product”.

```{r}
recipe <- read_csv('recipe.csv')
```

```{r}
has_wheat <- ingredients %>% 
  left_join(food_item, by = c('food_item' = 'item')) %>% 
  group_by(recipe) %>% 
  filter(type == 'Wheat product') %>% 
  select %>% 
  pull()

ingredients %>% 
  left_join(food_item, by = c('food_item' = 'item')) %>% 
  filter(recipe %in% has_wheat) %>% 
  group_by(recipe) %>% 
  summarize(total_cal = sum(`weight (oz)`*calories))  
```


## Question 2

1. Find names.txt and write regex patterns to answer following questions

a. Find all usernames that contain at least one numeric character.

```{r}
names <- read.table(file = 'names.txt', sep = '\n')
names <- names$V1
names[str_detect(names, pat_1_a)]
```

b. Find all usernames that are exactly four characters long and consist only of alphabetic

```{r}
names[str_detect(names, pat_1_b)]
```

c. Find all usernames following the conventional way of name format, i.e., the “given name” goes first, and the “family” name last, with any other names in-between. The names are separated by a single white space and each name should be uppercase letter followed by one or more lowercase letters.

```{r}
names[str_detect(names, pat_1_c)]
```

2. Find cards.txt and write the regex patterns to identify all the valid card numbers with the formatting rules below

a. Write a regex pattern to match valid Master card number and print all the valid
numbers, grouped into sets of 4 separated by a single space.

```{r}
cards <- read.table(file = 'cards.txt', sep = '\n')
cards <- cards$V1
cards[str_detect(cards, pat_2_a)]
```

b. Write a regex pattern to match valid Visa card number and print all the valid numbers,
grouped into sets of 4 separated by a single space.
Note: You may use more than one stringer function for both questions above.

```{r}
cards[str_detect(cards, pat_2_b)]
```


3. Find the passwords.txt and write regex patterns to answer the following questions

a. Write a regex pattern to identify the passwords that satisfies the requirements below.


• Minimum 8 characters

• Must contain at least one letter

• Must contain at least one digit

```{r}
passwords <- read.table(file = 'passwords.txt', sep = '\n')
passwords <- passwords$V1
passwords[str_detect(passwords, pat_3_a)]
```

b. Write a regex pattern to identify the passwords that satisfies the requirements below.


• Minimum 8 characters

• Must contain at least one uppercase character

• Must contain at least one lowercase character

• Must contain at least one digit
```{r}
passwords[str_detect(passwords, pat_3_b)]
```

4. Please load wordlists.RData to R and write regular expression patterns which will match all of the values in x and none of the values in y.


a. Ranges

```{r}
load('wordlists.RData')
range_x <- unlist(wordlists$Ranges[1])
range_y <- unlist(wordlists$Ranges[2])
all(str_detect(range_x, pat_4_a))
any(str_detect(range_y, pat_4_a))
```

b. Backrefs
```{r}
backrefs_x <- unlist(wordlists$Backrefs[1])
backrefs_y <- unlist(wordlists$Backrefs[2])
all(str_detect(backrefs_x, pat_4_b))
any(str_detect(backrefs_y, pat_4_b))
```

c. Prime
```{r}
primes_x <- unlist(wordlists$Prime[1])
primes_y <- unlist(wordlists$Prime[2])
all(str_detect(primes_x, pat_4_c))
any(str_detect(primes_y, pat_4_c))
```



