---
title: "102a Discussion 10"
author: "Joshua Susanto"
date: "2023-03-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stats)
```

# Homework Requirements
1. For those main functions (those written in bold letters), please provide flowchart OR
algorithms sufficiently complete, clear, and concise enough to enable a person to accurately
implement the function in any programming language they are adept with using.
2. Write the function(s) which accurately implements the algorithm(s) as described or
requested.
3. Include the error-handling to ensure your function(s) work properly

# 1: Root-Finding Problem

(1) Write R code using the Bisection method to solve the following equations within an error of 10−8
\n\n
Our interval is going to start at a = -3 and b = -2 as f(a)f(b) < 0 which implies that our solution in within these two numbers


$$(b-a)*(0.5)^n\leq10^{-8}\\
(-2+3)*(0.5)^n\leq10^{-8}\\
\log_{0.5}(0.5)^n\leq\log_{0.5}10^{-8}\\
n\geq26.57542475$$


Hence we expect 27 iterations for this problem

(a) x^3 + 23 = 0
```{r}
f_parta <- function(x) x^3 + 23
f_parta(1)
# want our initial points to have a negative product
x_l <- -3
x_r <- -2
f_l <- f_parta(x_l)
f_r <- f_parta(x_r)
f_parta(x_l) * f_parta(x_r)
tol <- 10^-8

its <- 0
while(x_r - x_l > tol) {
  x_m <- (x_l + x_r) / 2
  f_m <- f_parta(x_m)
  if(identical(all.equal(f_m,0),TRUE)){break}
  if(f_l * f_m < 0){
    x_r <- x_m
  } else{
    x_l <- x_m
  }
  its <- its + 1
}
x_m
its
```


(b) x^x − 18 = 0

```{r}
g_parta <- function(x) x^x -18
g_parta(1)
# want our initial points to have a negative product
x_l <- 2
x_r <- 3
g_l <- g_parta(x_l)
g_r <- g_parta(x_r)
g_parta(x_l) * g_parta(x_r)
tol <- 10^-8

its <- 0
while(x_r - x_l > tol) {
  x_m <- (x_l + x_r) / 2
  g_m <- g_parta(x_m)
  if(identical(all.equal(g_m,0),TRUE)){break}
  if(g_l * g_m < 0){
    x_r <- x_m
  } else{
    x_l <- x_m
  }
  its <- its + 1
}
its
x_m
```

(c) e^−x^2 − 1/10 = 0

```{r}
h_parta <- function(x) exp(-x^2) - 0.1
h_parta(1)
# want our initial points to have a negative product
x_l <- 1
x_r <- 2
h_l <- h_parta(x_l)
h_r <- h_parta(x_r)
h_parta(x_l) * h_parta(x_r)
tol <- 10^-8

its <- 0
while(x_r - x_l > tol) {
  x_m <- (x_l + x_r) / 2
  h_m <- h_parta(x_m)
  if(identical(all.equal(h_m,0),TRUE)){break}
  if(h_l * h_m < 0){
    x_r <- x_m
  } else{
    x_l <- x_m
  }
  its <- its + 1
}
its
x_m
```


Note: For all three cases, please state your starting interval and the estimated number of
iterations required.


(2) Please write R code using the Fixed point iteration method to solve the equation (b). Explain your algorithm prior to the code. (explain which g you are using and why)
Note: You may answer and write code for above two questions directly on Rmd file.



In order to start our fixed point algorithm, we need to find our three candidates for g(x) like so:

$$f(x) = x^x - 18 \\
g_1(x) = x - f(x) = -x^x + x - 18\\
g_2(x) = f(x) + x = x^x + x + 18\\
\\ Let \; f(x) = 0;\; 0 = x^x - 18\\
18 = x^x\\
ln(18) = ln(x)^x\\
ln(18) = xln(x)\\
g_3(x) = x = \frac{ln(18)}{ln(x)}$$


Now that we have our three candidates we need to check our convergence criterion for each case (with interval [2,3]) in order to find which choice of g will be best.

$$|g_1^{'}(x)| = |-x^x(1+ln(x)) + 1|\leq|g_1^{'}(3)|\;for\;x\in[2,3]\\
|g_1^{'}(3)| = |-55.66| > 1,\;so \;g_1(x)\; diverges\\
|g_2^{'}(x)| = |x^x(1+ln(x)) + 1|\leq|g_2^{'}(3)|\;for\;x\in[2,3]\\
|g_2^{'}(3)| = |-57.66| > 1,\;so \;g_2(x)\; diverges\\$$


$$|g_3^{'}(x)| = |-\frac{ln(18)}{xln^2(x)}|\\
|g_3^{'}(3) = |-0.798| < 1, \;so \;g_3\;converges\;and\;will\;be\;used$$



Since 3 is within our range of attraction, we will use x = 3 as our starting point



```{r}
g <- function(x){
  log(18)/log(x)
}
x_1 <- 3
x_2 <- g(x_1)
i <- 0
tol <- 1e-8
while(abs(x_2 - x_1) > tol){
  x_1 <- x_2
  x_2 <- g(x_1)
  i <- i + 1
}
x_2
```


(3) Use the Newton’s method to compute the square root by using only simple arithmetic
operations1

(a) Please write a function, get_sqrt(), that computes the square root of an arbitrary
positive number. The function should take four arguments, the last three have default
values:

a a non-negative number to compute the square root of.
tol determines when you consider two iterates to be equal.
iter max gives the maximum number of iterations.
verbose determines if you want to display intermediate results


```{r}
#include docstring, error handling, and algorithm
get_sqrt <- function(a, tol = 10^-8, iter_max = 1000, verbose = FALSE) {
  f <- function(x){
    x*x - a
  }
  f_prime <- function(x){
    2*x
  }
  x <- a
  its <- 0
  while((abs(f(x)) > tol) & (its < iter_max)){
    x <- x - f(x) / f_prime(x)
    its <- its + 1
    if(verbose){
      print(x)
    }
  }
  x
}
```


(b) Give an example of your function by computing the square root of 5.

```{r}
get_sqrt(5)
get_sqrt(5, verbose = TRUE)
```


(c) How would the program change to compute an arbitrary, positive integer, root? Please
use calculus to determine an appropriate iterative equation for approximating this
arbitrary root.

This program's internal function would change to determine an arbitrary root. In particular, the power of x will correspond to whatever root we decide to do. 


$$Let\;f(x) = x^n-a \\
0 = x^n - a\\
x = \sqrt[n]{a}\\
0 = f(x_0) + f′(x_0)(x_1 − x_0)\\
0 = f(x_0) + f′(x_0)x_1 − f′(x_0)x_0\\
f′(x_0)x_1 = f′(x_0)x_0 − f(x_0)\\
x_1 = x_0 − \frac{f(x_0)}{f′(x_0)} \\
x_1= x_0 - \frac{x_0^n-a}{nx_0^{n-1}}$$


(d) Write an R function, get abroot(), like the one for square root which calculates the
arbitrary root of a number. I.e., the root argument is required for the function.

```{r}
#include docstring, error handling, and algorithm
get_abroot <- function(a, root, tol = 10^-8, iter_max = 1000, verbose = FALSE) {
  f <- function(x){
    x^root - a
  }
  f_prime <- function(x){
    root*x^(root-1)
  }
  x <- a
  its <- 0
  while((abs(f(x)) > tol) & (its < iter_max)){
    x <- x - f(x) / f_prime(x)
    its <- its + 1
    if(verbose){
      print(x)
    }
  }
  x
}
```


(e) Give an example of your function by computing the seventh root of 13.

```{r}
get_abroot(13,7)
get_abroot(13, 7, iter_max = 1, verbose = TRUE)
```


(f) Let ek = |xk − root(7)√13|, where k indicates the kth iteration. Please print from e1 to e4, and specify your findings

```{r}
results <- c(get_abroot(13, 7, iter_max = 1), get_abroot(13, 7, iter_max = 2), get_abroot(13, 7, iter_max = 3), get_abroot(13, 4, iter_max = 5))
abs(results - get_abroot(13,7))
```



2. This problem accounts for 40% of this assignment.
(a) Using calculus, find the minimum of the function f(x) = x n − nα log(x), n = 1, 2, ... and α > 0.

$$f(x) = x^n − n\alpha ln(x),\; n = 1, 2, ... \;and\;\alpha > 0 \\
f′(x)= nx^{n-1}-\frac{n\alpha}{x}\\
0 = nx^{n-1}-\frac{n\alpha}{x}\\
xnx^{n-1}= n\alpha\\
nx^{n}= n\alpha\\
x = \sqrt[n]\alpha$$



$$f′′(x) = n(n-1)x^{n-2}+n\alpha x^{-2}\\
f′′(\alpha) = n(n-1)\alpha^{n-2}+n\alpha \alpha^{-2} > 0 \; for \; n = 1, 2, ... \;and\;\alpha > 0\\ $$


Hence, by the second derivative test, the minimum of the function occus at x = α



(b) Using your formulas for the derivatives of f, write down Newton’s method to find the
minimum.


$$f(x) = x^n − n\alpha ln(x),\; n = 1, 2, ... \;and\;\alpha > 0 \\
f′(x)= nx^{n-1}-\frac{n\alpha}{x}\\
f′′(x) = n(n-1)x^{n-2}+n\alpha x^{-2}\\
f′′(x_k)x_{k+1} = f′(x_k)x_k − f′(x_k)\\
x_{k+1} = x_k − \frac{f′(x_k)}{f′′(x_k)}\\
x_{k+1} = x_k − \frac{nx_k^{n-1}-\frac{n\alpha}{x_k}}{n(n-1)x_k^{n-2}+n\alpha x_k^{-2}}$$


(c) Use R to write a function, get_min(), which takes the following arguments:
f an expression to minimize with respect to x
x an initial guess
... any additional variables needed by f
Choose an appropriate stopping rule and, using your function, return the minimum of
f for n = 2 and α = 3

```{r}
n <- 2
a <- 3
f <- x^n - n*a*log(x)
f_expr <- expression(x^n - n*a*log(x))
f_expr
eval(f_expr)
f_prime <- stats::D(f_expr, name = "x")
eval(f_prime)
f_expr <- expression(x^2 + exp(-x))
f_prime <- D(f_expr, name = 'x')
?D
?expression
```

```{r}
#include docstring, error handling, and algorithm
get_min <- function(f, x, n, a, tol = 10^-8, iter_max = 1000) {
  f_expr <- f
  f_prime <- D(f_expr, name = "x")
  f_dub_prime <- D(f_prime, name = "x")
  x <- x
  its <- 0
  while((abs(eval(f_prime)) > tol) & (its < iter_max)){
    x <- x - eval(f_prime)/eval(f_dub_prime)
    its <- its + 1
  }
  x
}

get_min(expression(x^n-n*a*log(x)), x = 1, n = 2, a = 3)
typeof(1)
```

(d) Plot the function and the steps taken by your method starting from x0 = 1

