---
title: "Homework 1"
author: "Joshua Susanto"
date: '2023-01-22'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("405568250_stats102a_hw1.R")
```

## Question 1

Please write a function, gcd(), which takes two arguments x, y and calculates their greatest
common divisor (GCD) using the Euclidean algorithm.

Make use of gcd() to write a function, lcm(), which takes a vector and find the least common
multiple. The length of the vector will be at least two but no more than 100.

#### **`gcd()`**

**Algorithm:**

1. Check for errors and undesirable inputs.

2. Take the absolute values of x and y as GCD is not affected by sign

3. Initialize the larger input to be the dividend and the smaller value to be the divisor

4. Loop while the divisor is not equal to zero (as the Euclidean Algorithm will have our remainder be set as our new divisor)

5. Apply the Euclidean Algorithm by using the modulus operator to find the remainder, set the new dividend to be the old divisor and old divisor to be the remainder

6. Break from the loop when the remainder is equal to zero, return the current dividend
**Test Cases:**
```{r}
gcd(47011, 73618)  # expected output: 1
```
```{r error=TRUE}
gcd(11, '22')  # expected output: error
```
```{r}
gcd(-1, 531)  # expected output: 1
```
```{r}
gcd(72, 8)  # expected output: 8
```
```{r}
gcd(180,25)   # expected output: 5
```



#### **`lcm()`**

**Algorithm:**

1. Check for errors and undesirable inputs.

2. Check if length of x meets condition of being between 2 and 100

3. Initialize the first value to be the first lcm (lcm1)

4. Set another lcm (lcm2) to be the product of the first lcm and next value in the vector divided by their gcd (using the gcd function)

5. Set lcm1 to be lcm2

6. Repeat steps 4 and 5 until the final entry of the vector is reached. (Using a for loop)

7. Return lcm2

**Test Cases:**
```{r}
lcm(c(12, 21))  # expected output: 84
```
```{r error = TRUE}
lcm(c('2', 988))  # expected output: error
```
```{r}
lcm(c(4789, 6123, 199))  # expected output: 5835286353
```
```{r}
lcm(c(72, 8))  # expected output: 72
```
```{r error = TRUE}
lcm(c(4, 5, 0))   # expected output: error
```

## Question 2

Please write a function get_factors() which takes a number x and returns a list object. The list
object should contain the vector of unique prime factors of x and the corresponding exponents. In addition, write one “helper function” is_prime() that returns a logical vector depending on
whether or not the elements in x are prime.




For this problem I've defined another helper function check_prime() which takes a singular non-negative integer and returns a singular logical value indicating whether or not the integer is prime.

I've added algorithms as well as test cases for any helper functions defined as well.


#### **`check_prime()`**

**Algorithm:**

1. Check for errors and undesirable inputs. 

2. Initialize variable y to be 2 (smallest prime number)

3. See if y divides our input. If so, return false as it is automatically not prime.

4. If not, add 1 to y and keep repeating step 3 until y is no longer less than our input.

5. If y reaches our input, it tells us that our input is only divisible by itself. Thus, return TRUE as this means our input is prime

**Test Cases:**
```{r}
check_prime(92392)  # expected output: FALSE
```
```{r error = TRUE}
check_prime('22')  # expected output: error
```
```{r}
check_prime(73)  # expected output: TRUE
```
```{r}
check_prime(11)  # expected output: TRUE
```
```{r error = TRUE}
check_prime(0.88)   # expected output: error
```

#### **`is_prime()`** 

1. Set a variable n for the length of our vector and initialize an empty vector for the logical values of our output.

2. Establish a loop with length n

3. Check for errors and undesirable inputs, return an error or add FALSE to our output vector if necessary

4. Utilize helper function check_prime() in an if statement and add a TRUE into our logical vector if the condition is satisfied and FALSE if not.

5. After iterating through all entries of our input vector, return vector of logical values.

**Test Cases:**
```{r}
is_prime(c(9222, 727, 100, 231, 22))  # expected output: FALSE TRUE FALSE FALSE FALSE
```
```{r error = TRUE}
is_prime(c(9, 7, '1'))  # expected output: error
```
```{r}
is_prime(c(9, 7, 1))  # expected output: FALSE TRUE FALSE
```
```{r}
is_prime(c(72, 8))  # expected output: FALSE FALSE
```
```{r error = TRUE}
is_prime(c(9, 0, 1))   # expected output: error
```

#### **`get_factors()`** 

1. Check for errors and undesirable inputs and return an error if necessary.

2. Initialize an empty list for unique prime values and another for the corresponding exponents. Initialize variable y to be 2. (smallest prime number)

3. Check if y is prime, if so see if it divides our target number.

4. If the condition in step 3 is met, use another loop to count how many times y divides our input. Save y in our unique primes vector and the exponent count in the exponent vector. Update x to account for the division(s) by y. If the condition is not met, continue.

5. Add 1 to y and repeat steps 3 and 4 until x is 1, then return the factor and exponents vector as a list.

**Test Cases:**
```{r}
get_factors(1920)  # expected output: $primes c(2, 3, 5) $exponents c(7, 1, 1)
```
```{r error = TRUE}
get_factors(-28)  # expected output: error (ambiguity in negatives)
```
```{r}
get_factors(12)  # expected output: $primes c(2, 3) $exponents c(2, 1)
```
```{r}
get_factors(234)    # expected output: $primes c(2, 3, 13) $exponents c(1, 2, 1)
```
```{r}
# expected output: TRUE
x <- sample(x = 1e4, size = 1)
y <- get_factors(x)
this_works <- prod(y$primes ^ y$exponents) == x & all(is_prime(y$primes)) 
this_works
```

