---
title: "Stats 102a Homework 4"
output: html_document
date: "2023-03-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("405568250_stats102a_hw4.R")
```

## Question 1

(a) Write a constructor function, an appropriate predicate function, appropriate coercion
functions, and a useful print() method.

#### **`pqnumber()`**

**Test Cases:**
```{r}
ex1 <- pqnumber(1, 2, 0, c(4, 1, 3))
pqnumber(1, 2, 0, c(4, 1, 3))
```
```{r error=TRUE}
pqnumber(0.5, 2, 0, c(4, 1, 3)) # expected output: error
```
```{r}
ex2 <- pqnumber(sign = 1, p = 3, q = 4, nums = 1:8)
pqnumber(sign = 1, p = 3, q = 4, nums = 1:8)
```
```{r}
ex3 <- pqnumber(sign = -1, p = 5, q = 1, nums = c(2, 8, 2, 8, 1, 7, 2))
pqnumber(sign = -1, p = 5, q = 1, nums = c(2, 8, 2, 8, 1, 7, 2)) 
```
```{r}
ex4 <- pqnumber(sign = 1, p = 6, q = 0, nums = c(3, 9, 5, 1, 4, 1, 3))
pqnumber(sign = 1, p = 6, q = 0, nums = c(3, 9, 5, 1, 4, 1, 3)) 
```



#### **`is_pqnumber()`**

**Test Cases:**
```{r}
is_pqnumber(ex1) # expected output: true
```
```{r error=TRUE}
is_pqnumber(213) # expected output: false
```
```{r error=TRUE}
is_pqnumber(matrix(0,3,3)) # expected output: false
```
```{r}
is_pqnumber(ex3) # expected output: true 
```


#### **`print.pqnumber()`**

**Test Cases:**
```{r}
print.pqnumber(ex2)
```
```{r}
print.pqnumber(ex2, DEC = TRUE) # expected: 87654.321
```
```{r}
print.pqnumber(ex4)   
```
```{r}
print(ex3, DEC = TRUE) # expected: -27.18282  
```


#### **`as_pqnumber()`**

**Test Cases:**
```{r}
as_pqnumber(87654.321, 3, 4)
```
```{r}
as_pqnumber(3.1415926, 7, 0) 
```
```{r}
as_pqnumber.numeric(-27.18282, 5, 1)   
```
```{r error=TRUE}
as_pqnumber(123, 2, '2') # expected: error  
```



#### **`as_numeric.pqnumber()`**

**Test Cases:**
```{r}
as_numeric(ex1) # expected: 3.14
```
```{r}
as_numeric.pqnumber(ex2) # expected: 87654.321
```
```{r}
as_numeric.pqnumber(ex4) # expected: 3.141593
```
```{r}
as_numeric(ex3) # expected: -27.18282
```
```{r error=TRUE}
as_numeric.pqnumber(23) # expected: error
```



(b) Write an addition function add(x, y) and a subtraction function subtract(x, y).

My function utilizes abs_gtr() instead of borrow.

#### **`abs_gtr()`**

**Algorithm:**

1. Error handle as neccessary

2. Check whether a or b has the larger q value

3. If a has a larger q value, this means that there are more digits to the left of the decimal in a, since we want this value to be constant, we will fill the nums of a with 0 until the q values are the same

4. a is replaces with a new pqnumber that has the same q as b (extra spaces were filled with 0)

5. If b has a larger q value, then we do steps 3 and 4 with b

6. Check the the reverse elements of the nums of a and b, the instance one element is larger than the other, that pqnumber has a greater magnitude

7. If the first inputted pqnumber has a greater absolute value return TRUE and FALSE if otherwise


**Test Cases:**
```{r}
test1 <-  pqnumber(1, 3, 4, c(1,2,3,4, 0,0,0,0))
test2 <-  pqnumber(1, 3, 3, c(1,2,3,0, 1,0,0))
abs_gtr(test1, test2) # expected: false
```
```{r}
abs_gtr(test2, test1) # expected: true
```
```{r}
abs_gtr(ex2, ex1) # expected: true
```
```{r error=TRUE}
abs_gtr(ex3, 12) # expected: error
```
```{r}
abs_gtr(ex4, ex3) # expected: False
```
```{r}
abs_gtr(ex3, ex4) # expected: true
```



#### **`carry_over()`**

**Algorithm:**

1. Error handle as neccessary

2. Set a length n to the length of our vector

3. Set our initial 'carry' to be 0

4. Loop over the indices of our vector (utilizing n in our loop definition), set a value to be the sum of our carry and our value of our vector at the index

5. Mod the sum by 10 and append this value to our output vector, set our new carry to be the integer division of the sum by 10

6. repeat steps 4 and 5 for the rest of the vector and return output vector



#### **`add()`**

**Algorithm:**

1. Error handle as neccessary

2. Set our max p and q to be the sum of the ps and qs of x and y respectively

3. Set our length n to be the sum og our max p and q plus 1

4. set a vector z to be a vector of all zeroes of length n

5. Check if we have differing signs

6. If our signs are equal, then we can keep our x, y, and sign values as is

7. If our signs are different, we use abs_gtr() to check which pqnumber has the larger magnitude. If x is larger then we set our y values to be negative and our sign to be the sign of x. If y has a larger magnitude then we do the same thing except our x value are negative and we keep the sign of y.

8. With our established x values, y values, and sign, we add our x values and y values into z and use our carry_over function on z

9. We set our digit offset to be the length of z minus n in order to account for our final q value

10. We finally return a pqnumber with our set sign, max p, max q subtracted with our digit offset, and carried over z as our nums


**Test Cases:**
```{r}
print(add(ex1, ex2), DEC = TRUE) # 87657.461
```
```{r}
print(add(ex2, ex3), DEC = TRUE) # 87627.13818
```
```{r}
print(add(ex1, ex3), DEC = TRUE) # -24.04282
```
```{r}
print(add(ex3, ex1), DEC = TRUE) # -24.04282
```
```{r error=TRUE}
add(22, ex3) # error
```



#### **`subtract()`**

**Algorithm:**

1. Error handle as neccessary

2. Flip the sign of our second pqnumber (y)

3. Use add on our first number and our second flipped number

4. The nature of our add function ensures that the output will remain consistent due to our utilization of abs_gtr and carry_over

**Test Cases:**
```{r}
print(subtract(ex1, ex2), DEC = TRUE) # -87651.181
```
```{r}
print(subtract(ex2, ex3), DEC = TRUE) # 87681.50382
```
```{r}
print(subtract(ex1, ex3), DEC = TRUE) #30.32282
```
```{r}
print(subtract(ex3, ex1), DEC = TRUE) # -30.32282
```
```{r error=TRUE}
subtract(22, ex3) # error
```


(c) Multiplication
Use your add() function to write a multiplication function multiply(x, y) which can
multiply two pqnumber objects x and y. Think about how you would multiply two large
numbers by hand and implement that algorithm in R for two pqnumber objects. The function
should also return a pqnumber object. Both functions should return a pqnumber object with
enough p and q to carry the result. This question accounts for 15% of this assignment.

**Using Add not required**


#### **`multiply()`**

**Algorithm:**

1. Error handle as necessary

2. Set length n to be the sum of all ps and qs of x and y plus 1

3. Set vector z to be a 0 vector of length n

4. Iterate over the total length of y plus 1 (call iterator i)

5. Set our x leftover value to be the length of x, set the ith through ith + leftover elements of z to be the sum of our previous values of z (of these same indexed elements) with the product of numbers of x and the ith number of y

6. Repeat step 5 over the whole loop

7. carry over z using the carry_over() function

8. Set our digit offset to be the length of our carried over z minus n (total length)

9. Set our sign to be the product of the sign of x and y

10. Set our new p to be the sum of x's and y's p

11. Set our new q to be the length of our carried over z minus our new p minus 1

12. return our product using pqnumber() with our new sign, new p, new q, and carried over z as our nums


**Test Cases:**
```{r}
print(multiply(ex1, ex2), DEC = TRUE) # 275234.56794
```
```{r}
print(multiply(ex2, ex3), DEC = TRUE) # -2382691.62996522
```
```{r}
print(multiply(ex1, ex3), DEC = TRUE) # -85.3540548
```
```{r}
print(multiply(ex1, ex1), DEC = TRUE) # 9.8596
```
```{r error=TRUE}
subtract(22, ex3) # error
```